Reproducibility of published meta-analyses on clinical psychological interventions

### 2023-08-10 submission.Rmd
Accepted version

author:
 - name: "Rubén López-Nicolás"
   affiliation: "1"
   corresponding: yes
   address: "Facultad de Psicología, Campus de Espinardo, Universidad de Murcia, edificio n° 31, 30100, Murcia, Spain"
   email: "rlopez@um.es"
 - name: "Daniel Lakens"
   affiliation: "2"
 - name: "Jose A. López-López"
   affiliation: "1"
 - name: "Maria Rubio-Aparicio"
   affiliation: "3"
 - name: "Alejandro Sandoval-Lentisco"
   affiliation: "1"
 - name: "Carmen López-Ibáñez"
   affiliation: "1"
 - name: "Desirée Blázquez-Rincón"
   affiliation: "1"
 - name: "Julio Sánchez-Meca"
   affiliation: "1"

affiliation: 
 - id: "1"
   institution: "University of Murcia, Spain"
 - id: "2"
   institution: "Eindhoven University of Technology, The Netherlands"
 - id: "3"
   institution: "University of Alicante, Spain"

abstract: |
  Meta-analysis is one of the most useful research approaches, the relevance of which relies on its credibility. Reproducibility of scientific results could be considered as the minimal threshold of this credibility. We assessed the reproducibility of a sample of meta-analyses published between 2000-2020. From a random sample of 100 papers reporting results of meta-analyses of interventions in clinical psychology, 217 meta-analyses were selected. We first tried to retrieve the original data by recovering a data file, recoding the data from document files, or requesting it from original authors. Second, through a multi-stage workflow, we tried to reproduce the main results of each meta-analysis. The original data were retrieved for 67% (146/217) meta-analyses. \textcolor{blue}{While this rate showed an improvement over the years, in only 5\% of these cases was it possible to retrieve a data file ready for reuse.} Of these 146, 52 showed a discrepancy larger than 5% in the main results in the first stage. For 10 meta-analyses this discrepancy was solved after fixing a coding error \textcolor{blue}{of our data retrieval process} and for 15 of them it was considered approximately reproduced in a qualitative assessment. In the remaining meta-analyses (18%, 27/146), different issues were identified in an in-depth review, such as reporting inconsistencies, lack of data, or transcription errors. Nevertheless, the numerical discrepancies were mostly minor, with little or no impact on the conclusions. \textcolor{blue}{Overall, one of the biggest threats to the reproducibility of meta-analysis is related to data availability and current data sharing practices in meta-analysis.}

@@

## Purpose 
Previous research has revealed that there is room for improvement at different stages of the meta-analytic process pipeline. In this study our purpose is twofold. First, we broadened previous process reproducibility assessments by considering data availability on request and by contacting original authors to request required information to reproduce the meta-analysis. Second, we verified the outcome reproducibility of the meta-analyses that were process-reproducible using the available data. Where previous work focused on the reproducibility of primary effect sizes \textcolor{blue}{by recoding data from primary studies}, we explored meta-analysis outcome reproducibility using the primary \textcolor{blue}{data} already coded by the original authors. \textcolor{blue}{Therefore, we attempted to retrieve the data shared by the authors of the meta-analysis.}

@@

## Retrieval of primary data
In order to be able to reproduce meta-analyses of aggregate data, primary-level^[By primary-level data we mean aggregate data from included primary studies.] effects sizes and their associated standard errors are required. These are generally computed from statistics retrieved from the primary studies such as means, standard deviations or sample sizes. We attempted to retrieve the least processed data \textcolor{blue}{shared by the authors of the meta-analysis.} First, we sought for the statistics used to compute primary effect sizes (e.g., means, sd); second, we sought for the primary effects sizes already computed and their standard errors (or, alternatively, the sampling variances): finally, we sought for the primary effects sizes and their confidence limits, from which the standard errors were approximated as follows: 
$$se_i = (\frac{UB_i - LB_i}{2z_{\alpha/2}})$$
with $se_i$ being the standard error of the ith effect size, $UB_i$ and $LB_i$ the upper and lower confidence limits of confidence interval for the ith effect size, and $z_{\alpha/2}$ the $(1 - {\alpha/2})\%$ percentile of the standard normal distribution (usually, $z_{\alpha/2} = 1.96$ assuming a two-sided 95% confidence interval).

@@

## Process reproducibility 
Figure 3 summarizes the primary data retrieval results. Based on the availability of primary data, either retrieved directly from the paper or upon request, `r df_process_rep$n[2]` meta-analyses (`r sprintf("%0.0f%%", df_process_rep$perc[2]*100)`, see Fig. 3a) were labelled as process reproducible. \textcolor{blue}{Additionally, as the time span covered is fairly wide, the process reproducible rate was also computed for different time periods. The meta-analyses were grouped into five-year periods, except for the initial period, which was grouped into a ten-year period due to the limited number of meta-analyses available during the first five-year period, which consisted of only five meta-analyses. The process reproducibility rate was }`r sprintf("%0.0f%%, (%0.0f/%0.0f)", df_process_rep2$perc[2]*100, df_process_rep2$n[2], sum(df_process_rep2$n))`, `r sprintf("%0.0f%%, (%0.0f/%0.0f)", df_process_rep3$perc[2]*100, df_process_rep3$n[2], sum(df_process_rep3$n))`, and `r sprintf("%0.0f%%, (%0.0f/%0.0f)", df_process_rep4$perc[2]*100, df_process_rep4$n[2], sum(df_process_rep4$n))` \textcolor{blue}{for meta-analyses published between 2000 and 2010, 2011 and 2015, and 2016 and 2020, respectively (see Fig. 3a). This trend is further explored in the Supplementary file available at: https://osf.io/fjhpw.}

@@

Nevertheless, a more positive sign comes from the positive association between publication year and the possibility of retrieving the data. \textcolor{blue}{The results tentatively suggest a trend of improving data availability over the years, with a notable rate of 80\% observed in meta-analyses published between 2016 and 2020.} This observation could be related to the existence of well-established meta-analysis reporting guidelines. For instance, the first PRISMA guideline [@moher2009] encouraged meta-analyst to report results of primary studies (e.g. primary effect sizes and their confidence interval through a forest plot, as was a common scenario among the cases included in this project), and the latest PRISMA guideline [@page2021] which puts more emphasis on appropriate data sharing through data files ready for reuse. At the same time in only 5% of the cases where data were retrieved in our sample were we able to retrieve the data in a machine-readable data file that was ready for reuse (e.g., *csv*, *xlsx*). Most often the data had to be retrieved from files in document format (e.g., *docx*, *pdf*). This forces people who want to reuse the data to manually recode the data, which is an inefficient and error-prone task. Even after partial double coding was carried out, this procedure did not avoid some coding errors, which were only detected by double checking meta-analyses with discrepancies. In our experience, the data retrieval process can be difficult when results are presented in general tables, as it involves matching subsets of these primary data with different meta-analytic results, while is not always clear which studies were used in which meta-analysis reported in a paper. Furthermore, because the tables in manuscript are often generated manually in document file formats (e.g. Word), we observed examples where this introduced another source of error. The foregoing discussion raises a key point about how time consuming the appraisal of meta-analytic reproducibility currently is, and how efficiency would be improved by having open access to the underlying meta-analytic data in data file formats ready for reuse. The latest PRISMA guidelines, along with some initiatives that promote appropriate data sharing [e.g. @wilkinson2016] have the potential to generate significant improvements in the re-use of meta-analytic data in the years ahead. In this regard, our results provide a useful baseline for future assessments.

>>>>

author:
 - name: "Rubén López-Nicolás"
   affiliation: "1"
   corresponding: yes
   address: "Facultad de Psicología, Campus de Espinardo, Universidad de Murcia, edificio n° 31, 30100, Murcia, Spain"
   email: "rlopez@um.es"
 - name: "Daniel Lakens"
   affiliation: "2"
 - name: "Jose A. López-López"
   affiliation: "1"
 - name: "Maria Rubio-Aparicio"
   affiliation: "1"
 - name: "Alejandro Sandoval-Lentisco"
   affiliation: "1"
 - name: "Carmen López-Ibáñez"
   affiliation: "1"
 - name: "Desirée Blázquez-Rincón"
   affiliation: "1"
 - name: "Julio Sánchez-Meca"
   affiliation: "1"

affiliation: 
 - id: "1"
   institution: "University of Murcia, Spain"
 - id: "2"
   institution: "Eindhoven University of Technology, The Netherlands"

abstract: |
  Meta-analysis is one of the most useful research approaches, the relevance of which relies on its credibility. Reproducibility of scientific results could be considered as the minimal threshold of this credibility. We assessed the reproducibility of a sample of meta-analyses published between 2000-2020. From a random sample of 100 papers reporting results of meta-analyses of interventions in clinical psychology, 217 meta-analyses were selected. We first tried to retrieve the original data by recovering a data file, recoding the data from document files, or requesting it from original authors. Second, through a multi-stage workflow, we tried to reproduce the main results of each meta-analysis. The original data were retrieved for 67% (146/217) meta-analyses. While this rate showed an improvement over the years, in only 5\% of these cases was it possible to retrieve a data file ready for reuse. Of these 146, 52 showed a discrepancy larger than 5% in the main results in the first stage. For 10 meta-analyses this discrepancy was solved after fixing a coding error of our data retrieval process and for 15 of them it was considered approximately reproduced in a qualitative assessment. In the remaining meta-analyses (18%, 27/146), different issues were identified in an in-depth review, such as reporting inconsistencies, lack of data, or transcription errors. Nevertheless, the numerical discrepancies were mostly minor, with little or no impact on the conclusions. Overall, one of the biggest threats to the reproducibility of meta-analysis is related to data availability and current data sharing practices in meta-analysis.

@@

## Purpose 
Previous research has revealed that there is room for improvement at different stages of the meta-analytic process pipeline. In this study our purpose is twofold. First, we broadened previous process reproducibility assessments by considering data availability on request and by contacting original authors to request required information to reproduce the meta-analysis. Second, we verified the outcome reproducibility of the meta-analyses that were process-reproducible using the available data. Where previous work focused on the reproducibility of primary effect sizes by recoding data from primary studies, we explored meta-analysis outcome reproducibility using the primary data already coded by the original authors. Therefore, we attempted to retrieve the data shared by the authors of the meta-analysis.

@@

## Retrieval of primary data
In order to be able to reproduce meta-analyses of aggregate data, primary-level^[By primary-level data we mean aggregate data from included primary studies.] effects sizes and their associated standard errors are required. These are generally computed from statistics retrieved from the primary studies such as means, standard deviations or sample sizes. We attempted to retrieve the least processed data shared by the authors of the meta-analysis. First, we sought for the statistics used to compute primary effect sizes (e.g., means, sd); second, we sought for the primary effects sizes already computed and their standard errors (or, alternatively, the sampling variances): finally, we sought for the primary effects sizes and their confidence limits, from which the standard errors were approximated as follows: 
$$se_i = (\frac{UB_i - LB_i}{2z_{\alpha/2}})$$
with $se_i$ being the standard error of the ith effect size, $UB_i$ and $LB_i$ the upper and lower confidence limits of confidence interval for the ith effect size, and $z_{\alpha/2}$ the $(1 - {\alpha/2})\%$ percentile of the standard normal distribution (usually, $z_{\alpha/2} = 1.96$ assuming a two-sided 95% confidence interval).

@@

## Process reproducibility 
Figure 3 summarizes the primary data retrieval results. Based on the availability of primary data, either retrieved directly from the paper or upon request, `r df_process_rep$n[2]` meta-analyses (`r sprintf("%0.0f%%", df_process_rep$perc[2]*100)`, see Fig. 3a) were labelled as process reproducible. Additionally, as the time span covered is fairly wide, the process reproducible rate was also computed for different time periods. The meta-analyses were grouped into five-year periods, except for the initial period, which was grouped into a ten-year period due to the limited number of meta-analyses available during the first five-year period, which consisted of only five meta-analyses. The process reproducibility rate was `r sprintf("%0.0f%%, (%0.0f/%0.0f)", df_process_rep2$perc[2]*100, df_process_rep2$n[2], sum(df_process_rep2$n))`, `r sprintf("%0.0f%%, (%0.0f/%0.0f)", df_process_rep3$perc[2]*100, df_process_rep3$n[2], sum(df_process_rep3$n))`, and `r sprintf("%0.0f%%, (%0.0f/%0.0f)", df_process_rep4$perc[2]*100, df_process_rep4$n[2], sum(df_process_rep4$n))` for meta-analyses published between 2000 and 2010, 2011 and 2015, and 2016 and 2020, respectively (see Fig. 3a). This trend is further explored in the Supplementary file available at: https://osf.io/fjhpw.

@@

Nevertheless, a more positive sign comes from the positive association between publication year and the possibility of retrieving the data. The results tentatively suggest a trend of improving data availability over the years, with a notable rate of 80% observed in meta-analyses published between 2016 and 2020. This observation could be related to the existence of well-established meta-analysis reporting guidelines. For instance, the first PRISMA guideline [@moher2009] encouraged meta-analyst to report results of primary studies (e.g. primary effect sizes and their confidence interval through a forest plot, as was a common scenario among the cases included in this project), and the latest PRISMA guideline [@page2021] which puts more emphasis on appropriate data sharing through data files ready for reuse. At the same time in only 5% of the cases where data were retrieved in our sample were we able to retrieve the data in a machine-readable data file that was ready for reuse (e.g., *csv*, *xlsx*). Most often the data had to be retrieved from files in document format (e.g., *docx*, *pdf*). This forces people who want to reuse the data to manually recode the data, which is an inefficient and error-prone task. Even after partial double coding was carried out, this procedure did not avoid some coding errors, which were only detected by double checking meta-analyses with discrepancies. In our experience, the data retrieval process can be difficult when results are presented in general tables, as it involves matching subsets of these primary data with different meta-analytic results, while is not always clear which studies were used in which meta-analysis reported in a paper. Furthermore, because the tables in manuscript are often generated manually in document file formats (e.g. Word), we observed examples where this introduced another source of error. The foregoing discussion raises a key point about how time consuming the appraisal of meta-analytic reproducibility currently is, and how efficiency would be improved by having open access to the underlying meta-analytic data in data file formats ready for reuse. The latest PRISMA guidelines, along with some initiatives that promote appropriate data sharing [e.g. @wilkinson2016] have the potential to generate significant improvements in the re-use of meta-analytic data in the years ahead. In this regard, our results provide a useful baseline for future assessments.

### 2023-08-10 submission.tex
Accepted version

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{meta-analysis, reproducibility, data sharing, data reusability, research synthesis}
\usepackage{csquotes}
\usepackage{caption}
\captionsetup[table]{textfont={it}, labelfont={bf}, singlelinecheck=false, labelsep=newline}
\captionsetup[figure]{textfont={footnotesize, it}, labelfont={bf}, singlelinecheck=false, labelsep=newline}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Reproducibility of published meta-analyses on clinical psychological interventions},
  pdfauthor={Rubén López-Nicolás1, Daniel Lakens2, Jose A. López-López1, Maria Rubio-Aparicio3, Alejandro Sandoval-Lentisco1, Carmen López-Ibáñez1, Desirée Blázquez-Rincón1, \& Julio Sánchez-Meca1},
  pdflang={en-EN},
  pdfkeywords={meta-analysis, reproducibility, data sharing, data reusability, research synthesis},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Reproducibility of published meta-analyses on clinical psychological interventions}
\author{Rubén López-Nicolás\textsuperscript{1}, Daniel Lakens\textsuperscript{2}, Jose A. López-López\textsuperscript{1}, Maria Rubio-Aparicio\textsuperscript{3}, Alejandro Sandoval-Lentisco\textsuperscript{1}, Carmen López-Ibáñez\textsuperscript{1}, Desirée Blázquez-Rincón\textsuperscript{1}, \& Julio Sánchez-Meca\textsuperscript{1}}
\date{}

@@

\affiliation{\vspace{0.5cm}\textsuperscript{1} University of Murcia, Spain\\\textsuperscript{2} Eindhoven University of Technology, The Netherlands\\\textsuperscript{3} University of Alicante, Spain}

@@

\abstract{%
Meta-analysis is one of the most useful research approaches, the relevance of which relies on its credibility. Reproducibility of scientific results could be considered as the minimal threshold of this credibility. We assessed the reproducibility of a sample of meta-analyses published between 2000-2020. From a random sample of 100 papers reporting results of meta-analyses of interventions in clinical psychology, 217 meta-analyses were selected. We first tried to retrieve the original data by recovering a data file, recoding the data from document files, or requesting it from original authors. Second, through a multi-stage workflow, we tried to reproduce the main results of each meta-analysis. The original data were retrieved for 67\% (146/217) meta-analyses. \textcolor{blue}{While this rate showed an improvement over the years, in only 5\% of these cases was it possible to retrieve a data file ready for reuse.} Of these 146, 52 showed a discrepancy larger than 5\% in the main results in the first stage. For 10 meta-analyses this discrepancy was solved after fixing a coding error \textcolor{blue}{of our data retrieval process} and for 15 of them it was considered approximately reproduced in a qualitative assessment. In the remaining meta-analyses (18\%, 27/146), different issues were identified in an in-depth review, such as reporting inconsistencies, lack of data, or transcription errors. Nevertheless, the numerical discrepancies were mostly minor, with little or no impact on the conclusions. \textcolor{blue}{Overall, one of the biggest threats to the reproducibility of meta-analysis is related to data availability and current data sharing practices in meta-analysis.}
}

@@

Previous research has revealed that there is room for improvement at different stages of the meta-analytic process pipeline. In this study our purpose is twofold. First, we broadened previous process reproducibility assessments by considering data availability on request and by contacting original authors to request required information to reproduce the meta-analysis. Second, we verified the outcome reproducibility of the meta-analyses that were process-reproducible using the available data. Where previous work focused on the reproducibility of primary effect sizes \textcolor{blue}{by recoding data from primary studies}, we explored meta-analysis outcome reproducibility using the primary \textcolor{blue}{data} already coded by the original authors. \textcolor{blue}{Therefore, we attempted to retrieve the data shared by the authors of the meta-analysis.}

@@

In order to be able to reproduce meta-analyses of aggregate data, primary-level\footnote{By primary-level data we mean aggregate data from included primary studies.} effects sizes and their associated standard errors are required. These are generally computed from statistics retrieved from the primary studies such as means, standard deviations or sample sizes. We attempted to retrieve the least processed data \textcolor{blue}{shared by the authors of the meta-analysis.} First, we sought for the statistics used to compute primary effect sizes (e.g., means, sd); second, we sought for the primary effects sizes already computed and their standard errors (or, alternatively, the sampling variances): finally, we sought for the primary effects sizes and their confidence limits, from which the standard errors were approximated as follows:
\[se_i = (\frac{UB_i - LB_i}{2z_{\alpha/2}})\]
with \(se_i\) being the standard error of the ith effect size, \(UB_i\) and \(LB_i\) the upper and lower confidence limits of confidence interval for the ith effect size, and \(z_{\alpha/2}\) the \((1 - {\alpha/2})\%\) percentile of the standard normal distribution (usually, \(z_{\alpha/2} = 1.96\) assuming a two-sided 95\% confidence interval).

@@

Figure 3 summarizes the primary data retrieval results. Based on the availability of primary data, either retrieved directly from the paper or upon request, 146 meta-analyses (67\%, see Fig. 3a) were labelled as process reproducible. \textcolor{blue}{Additionally, as the time span covered is fairly wide, the process reproducible rate was also computed for different time periods. The meta-analyses were grouped into five-year periods, except for the initial period, which was grouped into a ten-year period due to the limited number of meta-analyses available during the first five-year period, which consisted of only five meta-analyses. The process reproducibility rate was }41\%, (12/29), 59\%, (44/75), and 80\%, (90/113) \textcolor{blue}{for meta-analyses published between 2000 and 2010, 2011 and 2015, and 2016 and 2020, respectively (see Fig. 3a). This trend is further explored in the Supplementary file available at: https://osf.io/fjhpw.}

Of these 146 meta-analyses, in about half of the cases the primary data was retrieved from a forest plot in the paper itself and in about a third of the cases the primary data was retrieved from supplementary files (see Fig. 3b for further details).
Although attempts were made to retrieve data for 75 meta-analyses from 25 different papers by emailing the corresponding authors, data was only retrieved for 4 meta-analyses, from 3 different papers (12\%, 3/25, see Fig. 3c). For the remaining 71 from 22 different papers, a reply providing some reasons not to share was received in 32\% (8/25, see Fig 3c), whereas no reply was received for the remainder of the meta-analyses. Table 1 summarises the different reasons corresponding authors given when data was not provided upon request.

@@

In most cases, when the meta-analytic data was available, it was shared in document formats. Data were retrieved from tables or forest plots in \emph{pdf} or \emph{docx} format---either in the document itself or in the supplementary materials---in 92\% (134/146) of the cases. This required a manual recoding of the primary data to be able to reuse them. Furthermore, when data was reported through general tables (i.e.~tables listing all the primary studies included with their characteristics), the meta-analysis associated with each data entry was not always obvious, leading to the time-consuming task of matching each data entry with each independent meta-analytic result reported in the paper. There were only 4 meta-analyses (from three different papers) of the 146 meta-analyses labelled as process reproducible (3\%), where the task of retrieving the data required simply downloading the data in an machine-readable data file format.
On the other hand, as shown in Figure 3c, when the necessary data was not available, retrieving it upon request to the original authors led to a low response rate.

@@

Nevertheless, a more positive sign comes from the positive association between publication year and the possibility of retrieving the data. \textcolor{blue}{The results tentatively suggest a trend of improving data availability over the years, with a notable rate of 80\% observed in meta-analyses published between 2016 and 2020.} This observation could be related to the existence of well-established meta-analysis reporting guidelines. For instance, the first PRISMA guideline (Moher et al., 2009) encouraged meta-analyst to report results of primary studies (e.g.~primary effect sizes and their confidence interval through a forest plot, as was a common scenario among the cases included in this project), and the latest PRISMA guideline (Page et al., 2021) which puts more emphasis on appropriate data sharing through data files ready for reuse. At the same time in only 5\% of the cases where data were retrieved in our sample were we able to retrieve the data in a machine-readable data file that was ready for reuse (e.g., \emph{csv}, \emph{xlsx}). Most often the data had to be retrieved from files in document format (e.g., \emph{docx}, \emph{pdf}). This forces people who want to reuse the data to manually recode the data, which is an inefficient and error-prone task. Even after partial double coding was carried out, this procedure did not avoid some coding errors, which were only detected by double checking meta-analyses with discrepancies. In our experience, the data retrieval process can be difficult when results are presented in general tables, as it involves matching subsets of these primary data with different meta-analytic results, while is not always clear which studies were used in which meta-analysis reported in a paper. Furthermore, because the tables in manuscript are often generated manually in document file formats (e.g.~Word), we observed examples where this introduced another source of error. The foregoing discussion raises a key point about how time consuming the appraisal of meta-analytic reproducibility currently is, and how efficiency would be improved by having open access to the underlying meta-analytic data in data file formats ready for reuse. The latest PRISMA guidelines, along with some initiatives that promote appropriate data sharing (e.g. Wilkinson et al., 2016) have the potential to generate significant improvements in the re-use of meta-analytic data in the years ahead. In this regard, our results provide a useful baseline for future assessments.

>>>>

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{meta-analysis, reproducibility, data sharing, data reusability, research synthesis}
\usepackage{csquotes}
\usepackage{caption}
\captionsetup[table]{textfont={it}, labelfont={bf}, singlelinecheck=false, labelsep=newline}
\captionsetup[figure]{textfont={footnotesize, it}, labelfont={bf}, singlelinecheck=false, labelsep=newline}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Reproducibility of published meta-analyses on clinical psychological interventions},
  pdfauthor={Rubén López-Nicolás1, Daniel Lakens2, Jose A. López-López1, Maria Rubio-Aparicio1, Alejandro Sandoval-Lentisco1, Carmen López-Ibáñez1, Desirée Blázquez-Rincón1, \& Julio Sánchez-Meca1},
  pdflang={en-EN},
  pdfkeywords={meta-analysis, reproducibility, data sharing, data reusability, research synthesis},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Reproducibility of published meta-analyses on clinical psychological interventions}
\author{Rubén López-Nicolás\textsuperscript{1}, Daniel Lakens\textsuperscript{2}, Jose A. López-López\textsuperscript{1}, Maria Rubio-Aparicio\textsuperscript{1}, Alejandro Sandoval-Lentisco\textsuperscript{1}, Carmen López-Ibáñez\textsuperscript{1}, Desirée Blázquez-Rincón\textsuperscript{1}, \& Julio Sánchez-Meca\textsuperscript{1}}
\date{}

@@

\affiliation{\vspace{0.5cm}\textsuperscript{1} University of Murcia, Spain\\\textsuperscript{2} Eindhoven University of Technology, The Netherlands}

@@

\abstract{%
Meta-analysis is one of the most useful research approaches, the relevance of which relies on its credibility. Reproducibility of scientific results could be considered as the minimal threshold of this credibility. We assessed the reproducibility of a sample of meta-analyses published between 2000-2020. From a random sample of 100 papers reporting results of meta-analyses of interventions in clinical psychology, 217 meta-analyses were selected. We first tried to retrieve the original data by recovering a data file, recoding the data from document files, or requesting it from original authors. Second, through a multi-stage workflow, we tried to reproduce the main results of each meta-analysis. The original data were retrieved for 67\% (146/217) meta-analyses. While this rate showed an improvement over the years, in only 5\% of these cases was it possible to retrieve a data file ready for reuse. Of these 146, 52 showed a discrepancy larger than 5\% in the main results in the first stage. For 10 meta-analyses this discrepancy was solved after fixing a coding error of our data retrieval process and for 15 of them it was considered approximately reproduced in a qualitative assessment. In the remaining meta-analyses (18\%, 27/146), different issues were identified in an in-depth review, such as reporting inconsistencies, lack of data, or transcription errors. Nevertheless, the numerical discrepancies were mostly minor, with little or no impact on the conclusions. Overall, one of the biggest threats to the reproducibility of meta-analysis is related to data availability and current data sharing practices in meta-analysis.
}

@@

Previous research has revealed that there is room for improvement at different stages of the meta-analytic process pipeline. In this study our purpose is twofold. First, we broadened previous process reproducibility assessments by considering data availability on request and by contacting original authors to request required information to reproduce the meta-analysis. Second, we verified the outcome reproducibility of the meta-analyses that were process-reproducible using the available data. Where previous work focused on the reproducibility of primary effect sizes by recoding data from primary studies, we explored meta-analysis outcome reproducibility using the primary data already coded by the original authors. Therefore, we attempted to retrieve the data shared by the authors of the meta-analysis.

@@

In order to be able to reproduce meta-analyses of aggregate data, primary-level\footnote{By primary-level data we mean aggregate data from included primary studies.} effects sizes and their associated standard errors are required. These are generally computed from statistics retrieved from the primary studies such as means, standard deviations or sample sizes. We attempted to retrieve the least processed data shared by the authors of the meta-analysis. First, we sought for the statistics used to compute primary effect sizes (e.g., means, sd); second, we sought for the primary effects sizes already computed and their standard errors (or, alternatively, the sampling variances): finally, we sought for the primary effects sizes and their confidence limits, from which the standard errors were approximated as follows:
\[se_i = (\frac{UB_i - LB_i}{2z_{\alpha/2}})\]
with \(se_i\) being the standard error of the ith effect size, \(UB_i\) and \(LB_i\) the upper and lower confidence limits of confidence interval for the ith effect size, and \(z_{\alpha/2}\) the \((1 - {\alpha/2})\%\) percentile of the standard normal distribution (usually, \(z_{\alpha/2} = 1.96\) assuming a two-sided 95\% confidence interval).

@@

Figure 3 summarizes the primary data retrieval results. Based on the availability of primary data, either retrieved directly from the paper or upon request, 146 meta-analyses (67\%, see Fig. 3a) were labelled as process reproducible. Additionally, as the time span covered is fairly wide, the process reproducible rate was also computed for different time periods. The meta-analyses were grouped into five-year periods, except for the initial period, which was grouped into a ten-year period due to the limited number of meta-analyses available during the first five-year period, which consisted of only five meta-analyses. The process reproducibility rate was 41\%, (12/29), 59\%, (44/75), and 80\%, (90/113) for meta-analyses published between 2000 and 2010, 2011 and 2015, and 2016 and 2020, respectively (see Fig. 3a). This trend is further explored in the Supplementary file available at: \url{https://osf.io/fjhpw}.

Of these 146 meta-analyses, in about half of the cases the primary data was retrieved from a forest plot in the paper itself and in about a third of the cases the primary data was retrieved from supplementary files (see Fig. 3b for further details).
Although attempts were made to retrieve data for 78 meta-analyses from 25 different papers by emailing the corresponding authors, data was only retrieved for 7 meta-analyses, from 3 different papers (12\%, 3/25, see Fig. 3c). For the remaining 71 from 22 different papers, a reply providing some reasons not to share was received in 32\% (8/25, see Fig 3c), whereas no reply was received for the remainder of the meta-analyses. Table 1 summarises the different reasons corresponding authors given when data was not provided upon request.

@@

In most cases, when the meta-analytic data was available, it was shared in document formats. Data were retrieved from tables or forest plots in \emph{pdf} or \emph{docx} format---either in the document itself or in the supplementary materials---in 92\% (134/146) of the cases. This required a manual recoding of the primary data to be able to reuse them. Furthermore, when data was reported through general tables (i.e.~tables listing all the primary studies included with their characteristics), the meta-analysis associated with each data entry was not always obvious, leading to the time-consuming task of matching each data entry with each independent meta-analytic result reported in the paper. There were only 7 meta-analyses (from three different papers) of the 146 meta-analyses labelled as process reproducible (5\%), where the task of retrieving the data required simply downloading the data in an machine-readable data file format.
On the other hand, as shown in Figure 3c, when the necessary data was not available, retrieving it upon request to the original authors led to a low response rate.

@@

Nevertheless, a more positive sign comes from the positive association between publication year and the possibility of retrieving the data. The results tentatively suggest a trend of improving data availability over the years, with a notable rate of 80\% observed in meta-analyses published between 2016 and 2020. This observation could be related to the existence of well-established meta-analysis reporting guidelines. For instance, the first PRISMA guideline (Moher et al., 2009) encouraged meta-analyst to report results of primary studies (e.g.~primary effect sizes and their confidence interval through a forest plot, as was a common scenario among the cases included in this project), and the latest PRISMA guideline (Page et al., 2021) which puts more emphasis on appropriate data sharing through data files ready for reuse. At the same time in only 5\% of the cases where data were retrieved in our sample were we able to retrieve the data in a machine-readable data file that was ready for reuse (e.g., \emph{csv}, \emph{xlsx}). Most often the data had to be retrieved from files in document format (e.g., \emph{docx}, \emph{pdf}). This forces people who want to reuse the data to manually recode the data, which is an inefficient and error-prone task. Even after partial double coding was carried out, this procedure did not avoid some coding errors, which were only detected by double checking meta-analyses with discrepancies. In our experience, the data retrieval process can be difficult when results are presented in general tables, as it involves matching subsets of these primary data with different meta-analytic results, while is not always clear which studies were used in which meta-analysis reported in a paper. Furthermore, because the tables in manuscript are often generated manually in document file formats (e.g.~Word), we observed examples where this introduced another source of error. The foregoing discussion raises a key point about how time consuming the appraisal of meta-analytic reproducibility currently is, and how efficiency would be improved by having open access to the underlying meta-analytic data in data file formats ready for reuse. The latest PRISMA guidelines, along with some initiatives that promote appropriate data sharing (e.g. Wilkinson et al., 2016) have the potential to generate significant improvements in the re-use of meta-analytic data in the years ahead. In this regard, our results provide a useful baseline for future assessments.
